---
title: "Homework 5 - Programming"
author: "Matthew Blevins"
date: "September 28, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
```
# Problem 1
I successfully completed the swirl lessons for this assignment.




# Problem 2
The file is created and in the correct place.





# Problem 3
There are several things that are important for showing data in a good figure.  First, it must be easy to read and understand without much explanation.  If it is too hard to understand, most people will not take the time to figure out what it represents.  It is also important for the correct type of figure to be selected depending on what kind of data is being portrayed.  There are many other important factors in creating a good figure, but many of them depend on the type of figure chosen.




# Problem 4
```{r, echo=F}
set.seed(432432)
#####Create a function to calculate the proportion of success
proportionOfSuccess <- function(vec){
#vec is the input vector
    sum(vec)/length(vec)
}

###Create a matrix to simulate 10 coin flips with varying degrees of fairness
p4bData <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10)


#Apply function and observe what happens
apply(p4bData, 2, proportionOfSuccess)
apply(p4bData, 1, proportionOfSuccess)

#The matrix just computed the overall probability when doing it by columns, and it gave me a vector that is essentially just the Probability data transposed when doing it by column.

#Fix the above matrix by creating a probability function and a 10 x 10 matrix
coinFlip <- function(p){
    rbinom(10, 1, prob = p)
}

probabilities <- c(.30, .31, .32, .33, .34, .35, .36, .37, .38, .39)

#Create a matrix with these probabilities
coinFlipMatrix <- matrix(sapply(probabilities, coinFlip), nrow = 10, ncol = 10)
coinFlipMatrix

#Apply by row
apply(coinFlipMatrix, 2, proportionOfSuccess)
#Apply by column
apply(coinFlipMatrix, 1, proportionOfSuccess)
```

#Problem 5
```{r, echo=F}
starchData <- read.table("http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat",
                          header=T)

#Begin by obtaining basic summary statistics of the data
summary(starchData)

#Plot a few boxplots to get an idea of the strength and thickness of the data
boxplot(strength ~ starch, starchData)

boxplot(thickness ~ starch, starchData)


#Plot the information together to gather information about all of the variables together
pairs(starchData)
cor(starchData$strength,starchData$thickness)

qplot(thickness, strength, data=starchData, col=starch)

g<-ggplot(starchData, aes(thickness, strength))
g+geom_point()+geom_smooth(method="lm")+facet_grid(.~starch)


```



# Problem 6


## Part A
```{r, echo=F}

 #we are grabbing a SQL set from here
    # http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

    #download the files, looks like it is a .zip
    library(downloader)
    download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",dest="us_cities_states.zip")
    unzip("us_cities_states.zip", exdir ="./05_R_apply_family")
    
    #read in data, looks like sql dump, blah
    library(data.table)
    states <- fread(input = "./us_cities_and_states/states.sql",skip = 23,sep = "'", sep2 = ",", header = F, select = c(2,4))
    ### YOU do the CITIES
    ### I suggest the cities_extended.sql may have everything you need
    ### can you figure out how to limit this to the 50?
    

#rename states columns 
colnames(states)<-c("state","Abbreviation") 

#read in cities 
cities <- fread(input = "./us_cities_and_states/cities_extended.sql",skip = 19,sep = "'", sep2 = ",", header = F, select = c(2,4,6,8,10,12)) 


#rename cities columns 
colnames(cities) <- c("city","Abbreviation","zip","latitude","longitude","county") 


#create vector of states 
abbreviation <- as.vector(unique(cities$Abbreviation)) 


#create vector of city counts by state 
citycount<-c()

for(i in 1:52){ 
    citycount[i] <- length(subset(cities$Abbreviation,cities$Abbreviation==abbreviation[i]))
}

knitr::kable(cbind.data.frame(abbreviation,citycount), caption="Number of Cities by State")


#####I ran out of time to finish the rest of this problem with having to study for the regression exam.  I'm planning to finish it Wednesday.
```
```


# Problem 7
If you are reading this, then the file was successfully pushed to github.