---
title: "HW3_Blevins_Matthew"
author: "Matthew Blevins"
date: "September 15, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(vioplot)
library(tidyverse)
library(data.table)
```
# Problem 1
I successfully completed the swirl lessons and installed the Exploratory Data Analysis courses.

# Problem 2
If you are reading this, then the file was successfully created and pushed to my github repository.

# Problem 3
There are several things that I took away from the two style guides from the lecture.  Firstly, it shows that there are still some style suggestions that are either controversial or being updated over time.  This means that one should note the style other collaborators have been using if they are jumping into the middle of a project.  The style guides also provide some nice tips for those who are just getting into coding to help keep things neat and tidy.  It can be intimidating to begin coding for the first time, so it certainly helps to maintain a uniform style that makes things easier to read.  Similarly to writing, everyone is going to develop their own style of coding, but there are still some basic rules that everyone should follow for simplicity's sake.  Some of the things that I will do specifically are to make sure that I am indenting appropriately so that the code is more readable, using appropriate spacing in my code, and avoiding the use of underscores in identifiers.  

# Problem 4
When using lintr, I received a lot of comments about spacing.  It told me to use spaces around my operators instead of typing them all together.  I had been writing them all together previously, but it does make the code more readable when there are spaces around them.


# Problem 5
``` {r, echo=F}
rds <- readRDS("HW3_data.rds")
rds.t <- arrange(rds, Observer)
dev1.means  <- data.frame()
dev2.means  <- data.frame()
dev1.sd     <- data.frame()
dev2.sd     <- data.frame()
dev12.cor   <- data.frame()

#Set a repeating loop for 1:13
for(i in 1:13){
    #Calculate the mean for each Observer using i
    mean1  <- mean(filter(rds, Observer==i)[, 'dev1'])
    mean2  <- mean(filter(rds, Observer==i)[, 'dev2'])
    
    #Calculate the standard deviations using i
    dev1sd    <-   sd(filter(rds, Observer==i)[, 'dev1'])
    dev2sd    <-   sd(filter(rds, Observer==i)[, 'dev2'])
    
    #Calculate the correlations
    correlate <-  cor(filter(rds, Observer==i)[, 'dev1'], 
                   filter(rds, Observer==i)[, 'dev2'])
    
    #Tie the calculations to a variable
    dev1.means  <- rbind(dev1.means, mean1)
    dev2.means <- rbind(dev2.means, mean2)
    
    dev1.sd     <- rbind(dev1.sd, dev1sd)
    dev2.sd     <- rbind(dev2.sd, dev2sd)
    
    dev12.cor <- rbind(dev12.cor, correlate)
}

#Create a table of all the data using knitr
devData <- cbind(1:13,dev1.means, dev2.means, dev1.sd, dev2.sd, dev12.cor)
colnames(devData) <- c('Observer','Dev1 Average','Dev2 Average', 'Dev1 SD', 'Dev2 SD', 'Correlation')
knitr::kable(devData, caption = "Means, Standard Dev., and Correlation for Devs.")


#Set up two boxplots beside each other
par(mfrow=c(1,2))
boxplot(dev1.means, main="Dev1 Averages")
boxplot(dev2.means, main="Dev2 Averages")

#Set up a violin plot
violinPlot <-  ggplot(devData, aes(x = dev1.sd, y= dev2.sd))
violinPlot + geom_violin()
```

# Problem 6
``` {r, echo=F}
library(dplyr)
bloodPressure <- read.table(
  'http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BloodPressure.dat', skip =2) 

#Change the names to something that makes sense regarding the data
colnames(bloodPressure)=c("Patient", "Morning1", "Afternoon1", "Evening1", "Patient", "Morning2", "Afternoon2", "Evening2")

#Combine second set of data to the first set
bloodPressure <- as.tibble(bloodPressure[, -5])
bloodPressureTidy <- cbind(c(bloodPressure$Morning1, bloodPressure$Afternoon1, bloodPressure$Evening1, bloodPressure$Morning2, bloodPressure$Afternoon2, bloodPressure$Evening2))


Patient <- as_tibble(rep(c(1:15), times=6))
measurementTime <- as_tibble(rep(c("Morning 1", "Afternoon 1", "Evening 1", "Morning 2", "Afternoon 2", "Evening 2"), each = 15))

bloodPressureFinal <- cbind(Patient, measurementTime, bloodPressureTidy)
colnames(bloodPressureFinal) <- c("Patient", "Measurement Time", "Blood Pressure")


invisible(as_tibble(bloodPressureFinal))
invisible(bloodPressureByPatient <- arrange(bloodPressureFinal, Patient))
knitr::kable(bloodPressureByPatient[1:90, ], caption = "Blood Pressure by Patient and Measurement Time")

```

# Problem 7
``` {r, echo=F}
newton <- function(f, tol=1E-18, x0=1, N=20) {
    h <- 0.001
    i <- 1
    x1 <- x0
    p <- numeric(N)
    
    while(i<=N) {
        df.dx <- (f(x0+h)-f(x0))/h
        x1 <- (x0 - (f(x0)/df.dx))
        p[i] <- x1
        i <- i + 1
        if (abs(x1-x0) < tol) break
        x0 <- x1
    }
            return(p[1:(i+1)])
}

f <- function(x){(3^x)-sin(x)+cos(x*5)}
b <- newton(f, x0=0, N=5)
b
```

# Problem 8
``` {r, echo=F}
library(data.table)
# Load the data

PersonData <- fread("C:/Users/Matthew Blevins/Documents/PersonData.csv")
GecData <- fread("C:/Users/Matthew Blevins/Documents/GecData.csv")
GebrekenData <- fread("C:/Users/Matthew Blevins/Documents/GebrekenData.csv")


#Merge the datasets, first by license plate then by defect code
    licensePlate <- merge(PersonData, GecData, by = "Kenteken")
    defectCode <- merge(licensePlate, GebrekenData, by = "Gebrek identificatie")
    
#Change the names of the columns to something I can understand
    defectsData <- defectCode[, c(1, 2, 4, 5, 16, 24)]
    colnames(defectsData) <- c("Defect Code", "License", "Make", "Model", "Inspection Date", "Defect Description")
    
#Check for NA
    checkForNA <- sum(is.na(defectsData$'Defect Code'))
    
#Examine only the year 2017 per the question
    defects2017 <- defectsData[grep("2017", defectsData$"Inspection Date"),]
    distinctMakes <- length(unique(defects2017$Make))
    distinctModels <- length(unique(defects2017$Model))
    
    knitr::kable(as.data.frame(cbind(distinctMakes, distinctModels)), caption = "Total Number of Distinct Makes and Models")
```

# Problem 9
This step is complete!
