---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

# Problem 1

If you are reading this, then the file has been created correctly.  :'( that there was no swirl problem.



# Problem 2
```{r, echo=F}
# Set the seed to maintain constant data set
#### I had to change the length.out for this to work on my computer.  It kept saying it couldn't perform the operations because the vector size was too large. 
set.seed(12345)
y <- seq(from = 1, to = 100, length.out = 1e+04) + rnorm(1e+04)
average <- mean(y)
    
#Create a for loop to calculate SST
SSTloop <- 0
for(i in 1:length(y)){
    SST1 <- SSTloop + (y[i] + average)^2
    SST1
}     



#Compute SST using vector operations
yPrime <- t(y)

#Create an n vector for identity matrix, then make a J matrix
n <- c(10000)
I <- matrix(0, nrow = n, ncol = n)
I[row(I)==col(I)] <- 1

J <- matrix(1, nrow = n, ncol = n)

#Create SST using matrix multiplication
SSTCen <- I - J/n
system.time({SST2 <- yPrime%*%SSTCen%*%y})
SST2
```


# Problem 3
```{r, echo=F}
set.seed(1256)
theta <- as.matrix(c(1, 2), nrow = 2)
X <- cbind(1, rep(1:10, 10))
h <- X %*% theta + rnorm(100, 0, 0.2)
m <- length(h)


oldTheta <- as.matrix(rnorm(2, 0, 1), nrow = 2)
alpha = 0.01
tol = 10e-6


#Run a loop to stop when the value stops changing within a reasonable limit
system.time({
while((abs(theta[1,1] - oldTheta[1,1]) & abs(theta[2,1] - oldTheta[2,1]))>tol) {   
    theta[1,1] <- theta[1,1] - alpha * 1/m * sum((theta[1,1] + theta[2,1] * X[,2] - h))
    theta[2,1] <- theta[2,1] - alpha * 1/m * sum((theta[1,1] + theta[2,1] * X[,2] - h) * X[,-2])
}
})


#Get the results for lm for comparison
system.time({
    lm(h~0+X)
})


```


# Problem 4

Instead of computing $(X'X)^{-1}$ It would be better to treat that inverse value as $A^{-1}$ and put it beside $\hat\beta$ as $A$. This would allow us to use the solve function without having to perform a large matrix multiplication, and it should be much faster that way.


# Problem 5
```{r, echo=F}
 set.seed(12456) 
    
    G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
    R <- cor(G) # R: 10 * 10 correlation matrix of G
    G <- NULL
    C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
    id <- sample(1:16000,size=932,replace=F)
    q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
    A <- C[id, -id] # matrix of dimension 932 * 15068
    B <- C[-id, -id] # matrix of dimension 15068 * 15068
    p <- runif(932,0,1)
    r <- runif(15068,0,1)
    C<-NULL #save some memory space
    
     #Calculate y
     system.time({
        y <- (p + (A%*%B^-1 %*% (q - r)))
        })
     ## My computer did not have enough memory to make the vectors, so I had to make them smaller to run the code and see what would work.  With the smaller vector, it took 55 seconds to compute this way, but it only took 10 seconds to solve it like I have below.
     
     system.time({
         t <- solve(B, q-r)
         y <- p + A %*% t
     })   
    
    
```

# Part a
A is 112,347,208 bytes
B is 1,816,357,192


# Part b



# Part c


# Problem 6
If you can see this, the pull request has been successfully completed.
