---
title: "HW2_Blevins_Matthew"
author: "Matthew Blevins"
date: "September 10, 2017"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(tidyr)
library(swirl)
```
# Problems 1-3

I successfully completed problems 1-3 to begin this assignment and create the file in the correct location.

# Problem 4

There are several valuable uses to utilizing version control in the classroom.  It allows others to work with existing code without getting rid of the original code.  This could be very useful in showing students what happens if certain things are changed.  It could also be very useful in working on group assignments because it provides a way for others to edit the code while still being able to revert to the original code if a bug is introduced.


# Problem 5

The code from problem 5 and 7 came from the solution posted on Dr. Settlage's github repository.  

``` {r, echo=F}
  #load the dataset
  #Sensory Table
url5a<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
Sensory_raw<-read.table(url5a, header=F, skip=1, fill=T, stringsAsFactors = F)
    Sensory_tidy<-Sensory_raw[-1,]
    Sensory_tidy_a<-filter(.data = Sensory_tidy,V1 %in% 1:10) %>%
                    rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
    Sensory_tidy_b<-filter(.data = Sensory_tidy,!(V1 %in% 1:10)) %>%
                    mutate(Item=rep(as.character(1:10),each=2)) %>%
                    mutate(V1=as.numeric(V1)) %>%
                    select(c(Item,V1:V5))
    Sensory_tidy<-bind_rows(Sensory_tidy_a,Sensory_tidy_b)
    colnames(Sensory_tidy)<-c("Item",paste("Person",1:5,sep="_"))
    Sensory_tidy<-Sensory_tidy %>%  
        gather(Person,value,Person_1:Person_5) %>%  
        mutate(Person = gsub("Person_","",Person)) %>%
        arrange(Item)
 knitr::kable(summary(Sensory_tidy), caption="Sensory data summary")

  #Long Jump Table
url5b<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
LongJump_raw<-read.table(url5b, header=F, skip=1, fill=T, stringsAsFactors = F)
    colnames(LongJump_raw)<-rep(c("V1","V2"),4)
    LongJump_tidy<-rbind(LongJump_raw[,1:2],LongJump_raw[,3:4],
                             LongJump_raw[,5:6],LongJump_raw[,7:8])
    LongJump_tidy<-LongJump_tidy %>%  
        filter(!(is.na(V1))) %>%
        mutate(YearCode=V1, Year=V1+1900, dist=V2) %>%
        select(-V1,-V2)
knitr::kable(summary(LongJump_tidy), caption="Long Jump data summary")
    
  #Brain and Body Weight Table
url5c<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
BrainBody_raw<-read.table(url5c, header=F, skip=1, fill=T, stringsAsFactors = F)
    colnames(BrainBody_raw)<-rep(c("Brain","Body"),3)
    BrainBody_tidy<-rbind(BrainBody_raw[,1:2],BrainBody_raw[,3:4],
                             BrainBody_raw[,5:6])
    BrainBody_tidy<-BrainBody_tidy %>%  
        filter(!(is.na(Brain)))
 knitr::kable(summary(BrainBody_tidy), caption="Brain/Body weight data summary")

 #Tomato Table
url5d<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
Tomato_raw<-read.table(url5d, header=F, skip=2, fill=T, stringsAsFactors = F, comment.char = "")
    Tomato_tidy<-Tomato_raw %>%  
        separate(V2,into=paste("C10000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
        separate(V3,into=paste("C20000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
        separate(V4,into=paste("C30000",1:3,sep="_"),sep=",",remove=T, extra="merge") %>%
        mutate(C10000_3=gsub(",","",C10000_3)) %>%
        gather(Clone,value,C10000_1:C30000_3) %>%
        mutate(Variety=V1, Clone=gsub("C","",Clone)) %>%
        mutate(Variety=gsub("\\\\#"," ",Variety)) %>%
        separate(Clone,into = c("Clone","Replicate")) %>%
        select(-V1,Variety,Clone,value) %>%
        arrange(Variety) 
knitr::kable(summary(Tomato_tidy), caption="Tomato data summary")

```

# Problem 6

``` {r, echo=F}
# Path to data
.datapath <- file.path(path.package('swirl'), 'Courses', 'R_Programming_E', 'Looking_at_Data','plant-data.txt')
# Read in data
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")

# Remove annoying columns
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% .cols2rm)]

# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period', 'Foliage_Color', 'pH_Min', 'pH_Max', 'Precip_Min', 'Precip_Max', 'Shade_Tolerance', 'Temp_Min_F')

knitr::kable(summary(plants), caption="Plants Tidy Data")

```


# Problem 7

``` {r, echo=F}
Car_Gebreken_raw <- file.path("C:", "Users", "Matthew Blevins", "Documents", "Open_Data_RDW__Gebreken.csv")
    Car_Geconstat_raw <-file.path("C:", "Users", "Matthew Blevins", "Documents", "Open_Data_RDW__Geconstateerde_Gebreken.csv")
    Car_Person_raw <- file.path("C:", "Users", "Matthew Blevins", "Documents","Personenauto_basisdata.csv")
    
    Car_Gebreken_raw.colclass <- sapply(Car_Gebreken_raw,class)
    Car_Geconstat_raw.colclass <- sapply(Car_Geconstat_raw,class)
    Car_Person_raw.colclass <- sapply(Car_Person_raw,class)
    
    print("Gebreken")
    print(Car_Gebreken_raw.colclass)

    print("Geconstat")
    print(Car_Geconstat_raw.colclass)

    print("Personen")
    print(Car_Person_raw.colclass)
```

